\subsection{Personal Notes}

\subsubsection{Aleksander Lundqvist}
I joined this group by chance, they had just lost a member and I was looking for a group during a break on a lecture. Now when we're about to reach the end of the project I fully realize that I had no idea what I was getting myself into. And I'm very happy to have joined this group and to have worked on this crazy project.

It's obvious that the workload of the project hasn't been equal between us. Henrik wrote the specification for the project, the signatures we used to create structures and the entire assembler, and has also been a great source of help when we've been stuck or even just not fully understood his specifications. Despite all that, we haven't been slacking off. This has been a huge project and has required quite a bit of studying up on things that have not been covered by the class, such as monads. When I finally got the monster of a function Vm.step to compile I felt triumphant. 130 lines of code. That was amazing. It didn't matter that I knew that it most likely was a buggy mess (this happened to be true), when it compiled I felt like we actually would be able to finish the project, even if not all features was implemented.

I have done most of my work by pair-programming with Oskar. During this time it has mostly been my hands on the keyboard, but having Oskar there to discuss with has been invaluable. And when I've been bugfixing he has contributed to the documentation, the part that is by far my weakest point in the project. 